---
title: "Problem Section 1"
subtitle:  "The z test"
graphics: yes
output: pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence. The specific tasks include: 

   - Perform a large sample z-test for the mean 
   
   - Perform a large sample z-test for a binomial proportion
   
   - Back up and support work with relevant explanations


\end{shaded}

* * *

### Exercises 

1. Suppose $n = 36$ observations are taken from a distribution with (unknown) mean $\mu_0$ and (known) standard deviation $\sigma_0 = 8.0$ for the purpose of testing
$H_0: \mu_0 = 60$ versus $H_1: \mu_0 \neq  60$ at the $\alpha = 0.07$ level of significance. (You may assume the CLT holds)

a. Find the rejection region of the $\alpha = 0.07$ level test. That is, what $\bar{x}$ values will the test reject the null hypothesis for?

b. Find the Type II error probability of the test from part a when $\mu_0 = 62$. 

c. Fill in the asterix in the expression for the confidence interval estimator which is the dual of the $\alpha = 0.07$ level significance test. Then find this number.
What is the confidence level of this interval?

    $$\bar{X} \pm z_{*} \frac{\sigma_0}{\sqrt{n} }$$
 

d. Suppose we observe the following data from this distribution. Calculate the P-value for the z-test of the hypotheses. What do you conclude at level $\alpha = 0.07$? Also, calculate your confidence interval from part c. Are the conclusions consistent?

    ```{r}

    x <- c( 60.38177, 55.85411, 56.00701, 40.08123, 57.13299, 59.62223,
        46.52059, 50.41849, 50.91220, 54.54830, 57.11061, 55.24721,
        73.26509, 61.65243, 64.88023, 45.96668, 57.81319, 55.62688,
        48.95113, 61.41349, 61.39330, 61.27571, 60.01789, 57.22339, 
        61.11239, 63.28603, 43.76773, 80.24444, 60.02996, 48.08963, 
        33.59426, 50.96914, 69.70207, 45.53624, 40.37203, 67.80395)
    ```



2. A manufacturer of rat poison claims that 82.87% of rats will be killed by their formulation. In order to verify the manufacturer's claim, an experiment is conducted where 150 rats are given
the poison, and 115 (76.67%) of them die. Do the results of this experiment indicate that the manufacturer has overestimated the effect of the poison? Let us conduct a level $\alpha  = 0.01$ significance test to find out.

    Suppose $X_i$ denote whether or not the $ith$ rat survives (0) or dies (1). Then we can assume
$$X_1,X_2,\dots,X_{150} \overset{i.i.d.}{\sim} Bernoulli(\pi_0)$$
    where $\pi_0$ is the true \emph{kill rate} of the rat poison.
    
    We are interested in testing the hypotheses in this model:
$$
H_0: \pi_0 = 0.8237 \ \ \ H_1: \pi_0 < 0.8237
$$

    
a. \textbf{Exact binomial test} Suppose the test will be based on the test statistic
   $$X = X_1 + X_2 + \dots + X_{150},$$
   the number of rats (out of 150) which will die when administered the poison.
   
    Answer the following questions.

    i. What is the distribution of $X$? 
    ii. Find the P-value to assess the evidence against $H_0$ and in favor of $H_1$ when we observe $x_{obs} = 115$.
    Hint: do large or small (or both) values of $X$ support $H_1$?
    iii. What do you conclude?

b. \textbf{z test} Suppose the test will be based on the test statistic
   $$\bar{X} = \frac{1}{150} \left(X_1 + X_2 + \dots + X_{150}\right)$$
   the fraction of rats (out of 150) which will die when administered the poison.

    Answer the following questions.

    i. What does the Central Limit Theorem say about the distribution of $\bar{X}$? 
    ii. Find the P-value to assess the evidence against $H_0$ and in favor of $H_1$ when we observe $\bar{x}_{obs} = \frac{115}{150} = `r round(115/150,4)`$.
    iii. What do you conclude?


